---
title: "Covid91 vaccine study Final2024"
author: "Johana Quintanilla"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "December 2025" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
males <- FinalData %>% filter(sex == "M")

ggplot(males, aes(x = treatment, fill = infected)) +
geom_bar(position = "dodge") +
labs(
title = "Vaccine Efficacy Among Males",
x = "Treatment Group",
y = "Number of Participants",
fill = "Infected"
)
```

This bar plot displays the number of infected and non-infected male participants in the treatment and control groups. The comparison highlights differences in infection outcomes between males who received the vaccine and those who did not. A lower number of infections in the treatment group would be consistent with evidence of vaccine efficacy among male participants. Among male participants, infection status differs between the treatment and control groups. A lower proportion of infected individuals in the treatment group would suggest that the vaccine is effective for males. Statistical testing is used to determine whether this observed difference is statistically significant.

## Females

```{r}
females <- FinalData %>% filter(sex == "F")

ggplot(females, aes(x = treatment, fill = infected)) +
geom_bar(position = "dodge") +
labs(
title = "Vaccine Efficacy Among Females",
x = "Treatment Group",
y = "Number of Participants",
fill = "Infected"
)
```

This bar plot compares infection status between treated and untreated female participants. Differences in the heights of the bars indicate variation in infection outcomes by treatment group. A smaller number of infected cases in the treated group would suggest that the vaccine reduces infection risk among females. For female participants, the comparison between treatment and control groups shows differences in infection outcomes. If the treated group exhibits fewer infections, this provides evidence that the vaccine is effective in reducing infection rates among females. The strength of this conclusion depends on the statistical significance of the results.

## LGBTQ

```{r}
lgbtq <- FinalData %>% filter(LGBTQ == "gay")

nrow(lgbtq)  # sanity check

ggplot(lgbtq, aes(x = treatment, fill = infected)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Vaccine Efficacy Among LGBTQ Participants",
    x = "Treatment Group",
    y = "Number of Participants",
    fill = "Infected"
  )

```

This bar plot displays the number of infected and non-infected participants who identify as gay across treatment and control groups. The figure allows for a visual comparison of infection outcomes within the LGBTQ subgroup. A lower number of infected individuals in the treatment group would suggest that the vaccine is effective in reducing infection risk among LGBTQ participants, though interpretation may be influenced by subgroup sample size. In this analysis, LGBTQ participants are defined as individuals who identify as gay, while straight-identifying participants are excluded from this subgroup.

## Druggies


# Overall Results and Conclusions