---
title: "Covid91 vaccine study Final2024"
author: "Johana Quintanilla"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "December 2025" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
males <- FinalData %>% filter(sex == "M")

ggplot(males, aes(x = treatment, fill = infected)) +
geom_bar(position = "dodge") +
labs(
title = "Vaccine Efficacy Among Males",
x = "Treatment Group",
y = "Number of Participants",
fill = "Infected"
)
```

This bar plot displays the number of infected and non-infected male participants in the treatment and control groups. The comparison highlights differences in infection outcomes between males who received the vaccine and those who did not. A lower number of infections in the treatment group would be consistent with evidence of vaccine efficacy among male participants. Among male participants, infection status differs between the treatment and control groups. A lower proportion of infected individuals in the treatment group would suggest that the vaccine is effective for males. Statistical testing is used to determine whether this observed difference is statistically significant.

## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions